---
title: "Task 3 Zvolen"
author: "Illia Shelepnytskyi"
date: "2025/11/23"
format:
  html:
    embed-resources: true
editor: source
---

###Task 3: Expanding Your Universe

Your “mini-universe” in Tasks 1 & 2 was a simulation. Now, it’s time to connect it to the real world.

Your mission is to find a real dataset that expands on your theme, import it into R, and explore it using the dplyr package.

3.1. The Quest: Find Your Data
Find one simple, real-world dataset that fits your theme.

Where to look: Kaggle, TidyTuesday, data.gov, or other public data portals.

Format: Any format you are able to read. For simplicity it is proposed to use textual (csv, txt) or Excel format.

Requirements:

At least 50 rows.

At least 4 columns.

A mix of numeric and character/factor data.

Action:

Download the data file and place it in your R project directory.

In your task file add a line with the link to the data source (e.g., Data from: https://…).

### 3.2. The Import: Load Your Data


Install tidyverse: If you haven’t, run install.packages("tidyverse") in your console (just once).

Load libraries: At the top of your script, add library(dplyr) and library(readr) (library(readxl) in case of Excel file format).

Read the data: Use read_csv() (or read_excel()) to load your dataset into a new data frame.

```{r}
# Load libraries
library(dplyr)
library(readr)

# Load data
zvolen_data <- read_csv("zvolen_df.csv")

# Quick inspection
glimpse(zvolen_data)
head(zvolen_data)
summary(zvolen_data)

```



###3.3. The Interrogation: Explore with dplyr
This is the core of the task. Your goal is to ask and answer 5 interesting questions about your new dataset.

You must write the question in plain English (as a comment), followed by the dplyr code that answers it.

You must use each of the following dplyr “verbs” at least once:

select() - To pick specific columns.

filter() - To subset rows based on a condition.

arrange() - To sort your results (e.g., with desc()).

mutate() - To create a new column.

group_by() and summarise() - To get summary statistics (like mean(), n(), max(), etc.).

Your 5 Questions (example):


**Question 1**: Top 5 streets by population?

```{r}
zvolen_data %>%
  select(district, street, population = pop) %>%
  arrange(desc(population)) %>%
  head(5)
```


**Question 2**: Average population by zone?

```{r}
zvolen_data %>%
  group_by(zone) %>%
  summarise(avg_population = mean(pop, na.rm = TRUE), n = n()) %>%
  arrange(desc(avg_population))
```


**Question 3**: Count of streets with a playground?

```{r}
zvolen_data %>%
  group_by(has_playground) %>%
  summarise(count = n())
```


**Question 4**: Population density of each street?

```{r}
zvolen_data %>%
  filter(!is.na(pop) & !is.na(area_km2)) %>%
  mutate(density = pop / area_km2) %>%
  select(district, street, density) %>%
  arrange(desc(density))
```


**Question 5**: Streets in the commercial zone?

```{r}
zvolen_data %>%
  filter(zone == "commercial") %>%
  select(district, street, pop, zone)
```