---
title: "Task 5 Zvolen"
author: "Illia Shelepnytskyi"
date: "2025/12/08"
format:
  html:
    embed-resources: true
editor: source
---

```{r setup, include=FALSE}
# Load required library
library(ggplot2)

# Load your dataset (make sure zvolen.csv is in the project folder)
zvolen_df <- read.csv("zvolen.csv")

# Ensure district is a factor for faceting
zvolen_df$district <- as.factor(zvolen_df$district)
```

## Task 5. Visualising Zvolen (Base R & ggplot2)

The goal is to create four visualizations for the dataset `zvolen_df`.

---

### 5.1. Quick Sketch (Base R Histogram)

Distribution of population across streets in Zvolen.

```{r histogram, echo=TRUE}
hist(zvolen_df$pop,
     main = "Distribution of Street Population in Zvolen",
     xlab = "Population (number of residents)",
     col = "lightblue",
     border = "white")
```

---

### 5.2. Relationship Map (ggplot2 Scatter Plot)

Relationship between street area and population.

```{r scatter, echo=TRUE}
ggplot(data = zvolen_df, aes(x = area_km2, y = pop)) +
  geom_point(alpha = 0.7, size = 3, colour = "darkblue") +
  geom_smooth() +
  labs(title = "Relationship between Street Area and Population – Zvolen",
       x = "Street Area (km²)",
       y = "Population") +
  theme_minimal()
```

---

### 5.3. Comparison (ggplot2 Boxplot)

Compare population by zone type (residential, commercial, industrial, mixed).

```{r boxplot, echo=TRUE}
ggplot(data = zvolen_df, aes(x = zone, y = pop, fill = zone)) +
  geom_boxplot() +
  labs(title = "Population Comparison by Zone in Zvolen",
       x = "Zone Type",
       y = "Population") +
  theme_light() +
  theme(legend.position = "none")
```

---

### 5.4. Deep Dive (Faceting)

Histogram of street areas faceted by district.

```{r facet, echo=TRUE}
ggplot(data = zvolen_df, aes(x = area_km2)) +
  geom_histogram(bins = 12, fill = "coral", color = "black") +
  facet_wrap(~ district) +
  labs(title = "Distribution of Street Areas by District – Zvolen",
       x = "Street Area (km²)",
       y = "Number of Streets")
```

---

### 5.5. Exporting (Saving Your Work)

If you want to save a separate PNG file:

```{r saveplot, echo=TRUE, eval=FALSE}
# Save the last plot as a PNG file
ggsave("zvolen_area_vs_pop.png", width = 8, height = 6)
```

> **Note:** For HTML output, `ggsave()` is optional because plots are embedded automatically.