---
title: "Task 4 Zvolen"
author: "Illia Shelepnytskyi"
date: "2025/12/1"
format:
  html:
    embed-resources: true
editor: source
---

## Task 4. Building a "Toolkit" for Your Universe (Functions)
In Task 3, you performed a "one-off" analysis of your new dataset. In the real world, you'll want to run that analysis again and again.

The best way to do this is by capturing your code inside **functions**. A function is a reusable "tool" that you build to perform a specific job.

Your mission is to build a "toolkit" of functions that can analyse the dataset you found in Task 3.

**Important:** You must use the **same dataset** and theme you used in Task 3.


### 4.1. The "Greeting" Function (A Simple Start)
Let's start with the basics. Write a simple function that prints a welcome message for your "universe".

1. **Write the function:** It should take at least one argument (e.g., `your_name` or `universe_name`).

2. **Inside the function:** Use `print()` or `cat()` to display a message.

3. **Call the function:** After you define the function, you must call it to see the result.

```{r}
welcome_to_my_universe <- function(your_name) {
  cat("--------------------------------------\n")
  cat(" Welcome to the Zvolen Analysis Toolkit\n")
  cat(" Report prepared by:", your_name, "\n")
  cat("--------------------------------------\n")
}

# Call the function:
welcome_to_my_universe("Illia Shelepnytskyi")

```

### 4.2. The "Loading" Function (A Good Practice)
In Task 3, you wrote code to load and inspect your data. Let's turn that into a function. This is a very common and useful practice.

1. **Write the function:** Name it `load_and_inspect`. It should take one argument: `file_path`.

2. **Inside the function:**

  - Load the `readr` and `dplyr` libraries.

  - Use `read_csv(file_path)` to load the data.

  - Print a message (e.g., "Data loaded successfully.").

  - Print the `glimpse()` of the data.

  - Print the `head()` of the data.

  - **Return** the data frame (use the `return()` command).

3. **Call the function:**

```{r}
# Load the libraries
library(readr)
library(dplyr)

# Define the function
load_and_inspect <- function(file_path) {
  
  # Load the data
  data <- read_csv(file_path)
  
  # Print messages for the user
  cat("--------------------------------------------------\n")
  cat(" Data loaded successfully from:", file_path, "\n")
  cat("--------------------------------------------------\n\n")
  
  cat("Glimpse of the dataset:\n")
  print(glimpse(data))
  
  cat("\nFirst 6 rows of the dataset:\n")
  print(head(data))
  
  # Return the loaded dataframe
  return(data)
}

# Call the function
zvolen_data <- load_and_inspect("zvolen_df.csv")

```


### 4.3. The "Query" Function (The Core of the Task)

```{r}
# Function: calculate population density for a selected district

population_density_in_district <- function(data, district_name) {
  
  result <- data %>%
    filter(district == district_name) %>% 
    filter(!is.na(pop) & !is.na(area_km2)) %>% 
    mutate(density = pop / area_km2) %>% 
    arrange(desc(density)) %>%
    select(district, street, density)
  
  return(result)
}

# Call the function
population_density_in_district(zvolen_data, "Zvolen-sever")
```

### 4.4. The "Reporting" Function

Let's combine everything. Write a single function that calculates several key statistics for a specific **group** or **category** in your data.

1. **Write the function:** Name it `generate_report`. It should take `data` and a `category_name` as arguments (e.g., a specific city, planet type, or region).

2. **Inside the function**:

    - `filter()` your data for only the `category_name` provided.

    - Calculate 3-4 interesting summary statistics (e.g., `mean()`, `max()`, `min()`, `n()`).

    - Use `cat()` to print a nicely formatted "report" to the console.

**Example:**


```{r}
generate_report <- function(data, zone_name) {
  
  zone_data <- data %>% 
    filter(zone == zone_name)
  
  # Calculate summary statistics
  count_records <- nrow(zone_data)
  avg_pop <- mean(zone_data$pop, na.rm = TRUE)
  max_pop <- max(zone_data$pop, na.rm = TRUE)
  avg_area <- mean(zone_data$area_km2, na.rm = TRUE)
  avg_density <- mean(zone_data$pop / zone_data$area_km2, na.rm = TRUE)
  
  # Print formatted report
  cat("\n----------------------------------------------\n")
  cat("      ZONE REPORT:", zone_name, "\n")
  cat("----------------------------------------------\n")
  cat("Number of Streets:", count_records, "\n")
  cat("Average Population:", round(avg_pop, 2), "\n")
  cat("Maximum Population:", max_pop, "\n")
  cat("Average Area (km²):", round(avg_area, 3), "\n")
  cat("Average Population Density:", round(avg_density, 2), "\n")
  cat("----------------------------------------------\n\n")
}

# Call the reporting function for two different zones:
generate_report(zvolen_data, "residential")
generate_report(zvolen_data, "commercial")

```
